---
title: "STACC Test - Task 2"
author: "Enlik Tjioe"
date: "2/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Plan:
- Setup the train and test data (filter all row, year < 2015)
- Setup the data to predict (all row with year == 2015)
- Build the ML Classification model

Reference:
file:///Users/enlik/GitRepo/_old/algoritma_ds_academy/07_ClassificationML01/C1/classification1.html#26_credit_risk_analysis__modeling:_loans_from_q4_2017



# Library
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(caret)

options(scipen = 9999)
```


## Data Pre-processing
```{r}
marketData <- read.csv("market_test_data.csv")
marketData$date <- dmy(marketData$date) # change data type of variable "date" from factor to Date
marketData <- select(marketData, -c("SP500")) # remove "SP500" variable 
str(marketData)
```

Filter data
```{r}
before2015 <- marketData %>% 
  filter(year(marketData$date) < 2015)
str(before2015)
```

```{r}
prop.table(table(before2015$label))
```

Testing LR model
```{r}
lr.model1 <- glm(label ~ usman.50, data = before2015, family = "binomial")
summary(lr.model1)
```

Setup data train and test
```{r}
set.seed(123)
intrain <- sample(nrow(before2015), nrow(before2015)*0.8)
before2015.train <- before2015[intrain, ]
before2015.test <- before2015[-intrain, ]
```

```{r}
lr.model2 <- glm(label ~ usman.50 + glocom.50 + usman.40 + glocom.40 + usman.growing + glocom.growing + usman.accel + glocom.accel + usman.3mgrowth	+ glocom.3mgrowth	+ profitgrowth.pos + revgrowth.pos + profitgrowth.accel + revgrowth.accel + industry.growing + industry.accel, data = before2015.train, family = "binomial")
summary(lr.model2)
```

```{r}
before2015.test$pred.label <- predict(lr.model2, before2015.test, type = "response")
before2015.test
```


```{r}
table("predicted"=as.numeric(before2015.test$pred.label>=0.5), "actual"=before2015.test$label)
```

```{r}
confusionMatrix(before2015.test$label, before2015.test$label)
```


```{r}
in2015 <- marketData %>% 
  filter(year(marketData$date) == 2015)
str(in2015)
```

```{r}
in2015$label <- predict(lr.model2, in2015, type="response")
in2015
```

```{r}
in2015$label = as.numeric(in2015$label>=0.5)
in2015
```

Export `in2015` dataframe to csv file
```{r}
write.csv(in2015, "labels_for_2015.csv", row.names = FALSE)
```

