---
title: "ML Capstone Sentiment Analysis"
author: "Enlik Tjioe"
date: "Updated: March 29, 2019"
output:
  html_document:
    css: style.css
    highlight: tango
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library
We will use `pacman` library for easier load/install packages with `p_load()` function
```{r}
library(pacman)
```

This tweets dataset is collected by Team Algoritma for educational purposes. In this project, we are going to analyze and build a classification model to predict sentiment for each tweets from “#YoutubeRewind2018”

The data is download by using twitter’s API. The data is available to load:

`fileEncoding="latin1"` was used to solve *Invalid multibyte* issue in relation with *file encoding*
# Pre-processing Data
```{r}
tweets <- read.csv("datasets/Sentiment Analysis/train.csv", fileEncoding="latin1")
head(tweets)
```

## Proportion per Sentiment Type
```{r}
prop.table(table(tweets$sentiment_type))
```
We can see that 40% of the tweets give *negative* sentiment, which the most one.


```{r}
smallData <- as.data.frame(prop.table(table(tweets$sentiment_type)))
colnames(smallData) <- c("Sentiment", "Proportion")
smallData
```

## Data Visualization
```{r}
p_load(ggplot2)
p_load(gridExtra)

# create blank theme for pie chart, otherwise it looks awful in my opinion
blank_theme = theme_minimal() + theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 14, face = 'bold') )

gbar = ggplot(smallData, aes(x = Sentiment, y = Proportion, fill = Sentiment))
gpie = ggplot(smallData, aes(x = "", y = Proportion, fill = Sentiment))

plot1 = gbar + geom_bar(stat = 'identity') + ggtitle("Overall Sentiment") + 
        theme(plot.title = element_text(size = 14, face = 'bold', vjust = 1),
              axis.title.y = element_text(vjust = 2), axis.title.x = element_text(vjust = -1))

plot2 = gpie + geom_bar(stat = 'identity') + coord_polar("y", start = 0) + blank_theme +
        theme(axis.title.x = element_blank()) + geom_text(aes(y = Proportion/3 + c(0, cumsum(Proportion)[-length(Proportion)]),
        label = round(Proportion, 2)), size = 4) + ggtitle('Overall Sentiment')

grid.arrange(plot1, plot2, ncol = 1, nrow = 2)
```

With data visualization, it's easier to see the proportion of sentiment

# Analysis of Tweets Text
```{r, warning=FALSE, message=FALSE}
p_load(tm) ; p_load(SnowballC) ; p_load(wordcloud)

# Custom function for transform corpus
transformer <- content_transformer(function(x, pattern){
  gsub(pattern, " ", x)
})
  
corpus <- Corpus(VectorSource(tweets$text))
  
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, transformer, "\\n")
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, removeWords, stopwords("english"))
corpus <- tm_map(corpus, stripWhitespace)
corpus <- tm_map(corpus, stemDocument)

tweets.dtm <- DocumentTermMatrix(corpus)

dim(tweets.dtm)
  
```

```{r, warning=FALSE, message=FALSE}
freqWords <- findFreqTerms(tweets.dtm, lowfreq = 3)
length(freqWords)
freqWords[1:10]
```

##  Visualize in Word Cloud
```{r, fig.width=8, fig.height=8}
wordcloud(corpus,
          min.freq = 5,
          max.words = 100,
          random.order = FALSE,
          colors = brewer.pal(5, "Set1"))
```



