---
title: "DV Capstone - Analysis FIFA 19 Player"
author: "Enlik Tjioe"
date: "4 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library
```{r}
library(tidyverse)
library(ggthemes)
library(lubridate)
library(rworldmap)
library(gplots)
library(knitr)
library(plotly)

options(scipen = 9999)
```

# Introduction
bla bla bla

**Some things I want to analyse:**

- Find out which project have the highest potential
- Find out relation between a and b (like in the book OUTLIER)

# Pre-processing Data
```{r}
ksdata <- read.csv("ks-projects-201801.csv")
str(ksdata)

```

```{r}
kable(head(ksdata))
```

```{r}
kable(tail(ksdata))
```

```{r}
sapply(ksdata, function(x) sum(is.na(x)))
```
We will remove column "usd.pledged" because it contains NA values, but we don't need it for Exploratory Data Analysis.

```{r}
ksdata <- ksdata[, -13]
colnames(ksdata)[13] <- "usd_pledged"
colnames(ksdata)[14] <- "usd_goal"
```


# What types of projects are most popular?
```{r}
cat.freq <- ksdata %>% 
  group_by(main_category) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

cat.freq$main_category <- factor(cat.freq$main_category, levels = cat.freq$main_category)
  
ggplot(cat.freq, aes(x = main_category, y = count)) +
  geom_bar(stat = "identity", aes(fill = main_category), show.legend = F) +
  labs(title = "Total Projects per Category", x = "Category Name", y = "Total") +
  theme(axis.text.x = element_text(angle=90, hjust=1),
        plot.title=element_text(hjust=0.5))
```

Film & Video is the most popular project category in Kickstarter


# What types of projects are being funded?
Top 20 highest funded projects
```{r}
kable(head(ksdata[order(-ksdata$usd_pledged), c(2,3,13)], 20))
```

Top 20 project with highest total backers
```{r}
kable(head(ksdata[order(-ksdata$backers), c(2,3,11)],20))
```


