---
title: "DV Capstone - Analysis FIFA 19 Player"
author: "Enlik Tjioe"
date: "4 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library
```{r}
library(tidyverse)
library(knitr)
library(plotly)

options(scipen = 9999)
```

# Introduction
bla bla bla

**Some things I want to analyse:**

- What types of projects were most popular?
- What types of projects were being funded?
- What types of projects were successful and unsuccessful?
- How many Successful and Failed Projects per Year?

# Pre-processing Data
```{r}
ksdata <- read.csv("ks-projects-201801.csv")
str(ksdata)

```

```{r}
kable(head(ksdata))
```

```{r}
kable(tail(ksdata))
```

```{r}
sapply(ksdata, function(x) sum(is.na(x)))
```
We will remove column "usd.pledged" because it contains NA values, but we don't need it for Exploratory Data Analysis.

```{r}
ksdata <- ksdata[, -13]
colnames(ksdata)[13] <- "usd_pledged"
colnames(ksdata)[14] <- "usd_goal"
```


# What types of projects were most popular?
```{r, fig.height=6, fig.width=10}
project.category <- ksdata %>% 
  group_by(main_category) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))

project.category$main_category <- factor(project.category$main_category, levels = project.category$main_category)
  
ggplot(project.category, aes(x = main_category, y = count)) +
  geom_bar(stat = "identity", aes(fill = main_category), show.legend = F) +
  labs(title = "Total Projects per Category", x = "Category Name", y = "Total") +
  theme(axis.text.x = element_text(angle=90, hjust=1),
        plot.title=element_text(hjust=0.5)) +
  geom_text(aes(label = paste0(round(count/1000, 1), "K")), vjust = -0.5)
```

Film & Video is the most popular project category in Kickstarter


# What types of projects were being funded?
Top 20 highest funded projects
```{r}
kable(head(ksdata[order(-ksdata$usd_pledged), c(2,3,13)], 20))
```

Top 20 project with highest total backers
```{r}
kable(head(ksdata[order(-ksdata$backers), c(2,3,11)],20))
```

Visualization
```{r, fig.height=6, fig.width=10}
category.pledged <- ksdata %>% 
  group_by(main_category) %>% 
  summarize(total = sum(usd_pledged)) %>% 
  arrange(desc(total))

category.pledged$main_category <- factor(category.pledged$main_category, levels = category.pledged$main_category)

ggplot(category.pledged, aes(x = main_category, y = total / 1000000)) +
  geom_bar(stat = "identity", aes(fill = main_category)) +
  labs(title = "Total Amount Pledged by Category", x = "Project Category", y = "Amount Pledged (in Millions USD)") +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 90, hjust = 1), legend.position="null") +
  geom_text(aes(label = paste0("$", round(total/1000000, 1))), vjust = -0.5)
  
```
Games, Design, and Technology are the highest grossing categories by far.



# What types of projects were successful and unsuccessful?

```{r, fig.height=6, fig.width=12}
state.freq <- ksdata %>% 
  group_by(state) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))

state.freq$state <- factor(state.freq$state, levels = state.freq$state)

ggplot(state.freq, aes(x = state, y = count)) +
  geom_bar(stat = "identity", aes(fill = state), show.legend = F) +
  ggtitle("Project by Status") + xlab("Status") + ylab("Total") +
  geom_text(aes(label = paste0(round(count/1000,1), "K")), vjust = -0.5)
```

In truth, failed projects was more than successful projects.

```{r}
state.grp <- ksdata %>%
  filter(state!="undefined") %>%
  mutate(grp=ifelse(state %in% c("successful", "failed"), "complete", "incomplete")) %>%
  group_by(grp, state) %>%
  summarize(count=n()) %>%
  mutate(pct=count/sum(count)) %>%
  arrange(grp, desc(-state))

state.grp$state <- factor(state.grp$state, levels=state.grp$state)

ggplot(state.grp, aes(grp, pct, fill=state)) + geom_bar(stat="identity") + 
  ggtitle("Project Status by Completion") + xlab("Project Completion") + ylab("Percentage") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")),
            position=position_stack(vjust=0.5), 
            colour="white", size=5) +
  theme(plot.title=element_text(hjust=0.5), 
        axis.title=element_text(size=12, face="bold"),
        axis.text.x=element_text(size=12), legend.position="bottom",
        legend.title=element_text(size=12, face="bold")) +
  scale_y_continuous(labels=scales::percent)
  
```

```{r fig.height=6, fig.width=12}
state.pct <- ksdata %>%
  filter(state %in% c("successful", "failed")) %>%
  group_by(main_category, state) %>%
  summarize(count=n()) %>%
  mutate(pct=count/sum(count)) %>%
  arrange(desc(state), pct)

state.pct$main_category <- factor(state.pct$main_category, 
                                  levels=state.pct$main_category[1:(nrow(state.pct)/2)])

ggplot(state.pct, aes(main_category, pct, fill=state)) + geom_bar(stat="identity") + 
  ggtitle("Success vs. Failure Rate by Project Category") + 
  xlab("Project Category") + ylab("Percentage") + scale_y_continuous(labels=scales::percent) + 
  scale_fill_discrete(name="Project Status", breaks=c("successful", "failed"),
                      labels=c("Success", "Failure")) + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), position=position_stack(vjust=0.5), 
            colour="white", size=5) + theme_economist() + 
  theme(plot.title=element_text(hjust=0.5), axis.title=element_text(size=12, face="bold"), 
        axis.text.x=element_text(size=12), legend.position="bottom", 
        legend.title=element_text(size=12, face="bold")) + coord_flip()
```

# How many Successful and Failed Projects per Year?
```{r fig.height=6, fig.width=12}
year.freq <- ksdata %>%
  filter(year(launched)!="1970") %>%
  group_by(year=year(launched)) %>%
  summarize(count=n())

ggplot(year.freq, aes(year, count, fill=count)) + geom_bar(stat="identity") + 
  ggtitle("Number of Projects by Launch Year") + xlab("Year") + ylab("Frequency") + 
  scale_x_discrete(limits=c(2009:2018)) + 
  geom_text(aes(label=paste0(count)), vjust=-0.5) + theme_economist() + 
  theme(plot.title=element_text(hjust=0.5), axis.title=element_text(size=12, face="bold"), 
        axis.text.x=element_text(size=12), legend.position="null")
```

```{r fig.height=6, fig.width=12}
state.pct2 <- ksdata %>%
  filter(year(launched)!="1970", state %in% c("successful", "failed")) %>%
  group_by(year=year(launched), state) %>%
  summarize(count=n()) %>%
  mutate(pct=count/sum(count)) %>%
  arrange(desc(state))

ggplot(state.pct2, aes(year, pct, fill=state)) + geom_bar(stat="identity") + 
  ggtitle("Success vs. Failure Rate by Year Launched") + 
  xlab("Year") + ylab("Percentage") + scale_x_discrete(limits=c(2009:2017)) + 
  scale_y_continuous(labels=scales::percent) + 
  scale_fill_discrete(name="Project Status", breaks=c("successful", "failed"),
                      labels=c("Success", "Failure")) + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), position=position_stack(vjust=0.5), 
            colour="white", size=5) + theme_economist() + 
  theme(plot.title=element_text(hjust=0.5), axis.title=element_text(size=12, face="bold"), 
        axis.text.x=element_text(size=12), legend.position="bottom", 
        legend.title=element_text(size=12, face="bold"))

```

