---
title: "Learn dplyr and plotly"
author: "Enlik Tjioe"
date: "28 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library
```{r}
library(dplyr)
library(plotly)
```

## Read Data
```{r}
vids <- read.csv("youtubetrends.csv")
str(vids)
```

## 1. dplyr
## 1.1 select
select based on column(s)
```{r}
vids.latihan <- select(vids, title, likes, dislikes, views)
head(vids.latihan)
```


## 1.2 filter
select based on row(s)
```{r}
vids.latihan <- filter(vids.latihan, views > 1000000 & dislikes < 100)
head(vids.latihan)
```

## 1.3.pipes (%>%)
make life easier
shortcut = CMD + SHIFT + M (Mac) or CTRL + SHIFT + M (Windows)
```{r}
vids.dplyr <- vids %>% 
                select(title, likes, views) %>% 
                filter(likes > 5000)

head(vids.dplyr)
```

## 1.4 mutate
make a new variable
```{r}
viral <- vids %>%
            mutate(viral = likes + dislikes + views)

head(vids)
```

## 1.5 group_by
Grouping data based on condition
```{r}
vids %>%
  group_by(category_id) %>% 
  filter(likes > 5000)
```

## 1.6 summarise
Summarise the dataframe, usually used with group_by
```{r}
vids %>%
  group_by(category_id) %>% 
  summarise(likesratio = sum(likes)/sum(views))
```

## 1.7 arrange
It sorting the dataframe
```{r}
vids %>% 
  group_by(category_id) %>% 
  summarise(likesratio = sum(likes) / sum(views)) %>% 
  arrange(desc(likesratio), desc(category_id))
```

## 1.8 Exercise
From data vids select variable title, category_id, views, likes, dislikes, and comment_count
with condition views > 500k and likes > 1000
and then create a new variable `viral` = (likes+dislikes) / views
then sort based on viral in descending values

```{r}
vids.kuis <- 
  vids %>%
    select(title, category_id, views, likes, dislikes, comment_count) %>%
    filter(views > 500000 & likes > 1000) %>%
    mutate(viral = (likes + dislikes) / views) %>% 
    arrange(desc(viral))

head(vids.kuis)
              
```


## 2. plotly
```{r}
theme_algoritma <- theme(legend.key = element_rect(fill="black"),
           legend.background = element_rect(color="white", fill="#263238"),
           plot.subtitle = element_text(size=6, color="white"),
           panel.background = element_rect(fill="#dddddd"),
           panel.border = element_rect(fill=NA),
           panel.grid.minor.x = element_blank(),
           panel.grid.major.x = element_blank(),
           panel.grid.major.y = element_line(color="darkgrey", linetype=2),
           panel.grid.minor.y = element_blank(),
           plot.background = element_rect(fill="#263238"),
           text = element_text(color="white"),
           axis.text = element_text(color="white")
           
           )
```



```{r}
# plotratio <- vids %>% 
#   mutate(likesratio = likes/views)
  
rangeslider()
```

