---
title: "LBB Regression Models"
author: "Enlik"
date: "14 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

**Learn-by-building Module**  
As this is a graded task for our Academy students, completion of the task is not optional and count towards your final score

Write a regression analysis report applying what you've learned in the workshop. Using the dataset provided by you, write your findings on the different socioeconomic variables most highly correlated to crime rates (`crime_rate`). Explain your recommendations where appropriate. To help you through the exercise, you should ask the following questions of your candidate model:  

- Can your model be any more simplified without losing substantial information?  
- Have you tried predicting the crime rate given a reasonable set of values for the predictor variable?  
- Have you identified any non-random pattern in your residual plot?

Students should be awarded the full points if:  
1. Student can select variable that is significant to both statistically and business-wise.
2. Student perform regression linear assumption checking (Linearity, Multicolinearity, Heteroscedasticity, Normality)
3. Student write the findings and explaining recommendations where appropriate.


Step by Step:
1. Read the data
2. Check the outlier
3. Check the correlation between Y ~ X, and every X
4. Create a linear model
5. Stepwise
6. Check other assumptions

# Library
```{r}
library(dplyr)
```

# Read the data
```{r}
crime <- read.csv("data_input/crime.csv") 
str(crime)

crime <- crime %>% 
 dplyr::select(-X)

names(crime) <- c("percent_m", "is_south", "mean_education", "police_exp60", "police_exp59", "labour_participation", "m_per1000f", "state_pop", "nonwhites_per1000", "unemploy_m24", "unemploy_m39", "gdp", "inequality", "prob_prison", "time_prison", "crime_rate")
plot(crime$gdp, crime$inequality)

```



```{r}

```

















