---
title: "20 Years of Video Games 1996-2016"
author: "Enlik Tjioe"
date: "27 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Library
```{r}
library(ggplot2)
```


## 2. Kaggle Dataset
[20 Years of Games](https://www.kaggle.com/egrinstein/20-years-of-games)

## 3. Background
In this data analysis, I will focusing to three big names in console game industry. They are Nintendo, Sony, and Microsoft. I will try to visualize who exactly win the **Console Wars** during 20 years timeline between 1996-2016

## 4. Pre-processing
```{r}
ign <- read.csv("data_input/ign.csv", stringsAsFactors = FALSE)
#str(ign)

ign[, c("score_phrase", "platform", "genre", "editors_choice")] <- lapply(ign[, c("score_phrase", "platform", "genre", "editors_choice")] , as.factor)

ign20 <- ign[ign$release_year >= 1996 & ign$release_year <= 2016, ]

## save memory by removing unused data frame
rm(ign)
```


## 5. Data Preparation
I define new variables for only 3 big console game vendor:
Nintendo, Sony, and Microsoft

```{r}
nintendo <- c("Game Boy","Game Boy Advance","Game Boy Color" ,"GameCube", "NES", "New Nintendo 3DS" ,"Nintendo 3DS","Nintendo 64", "Nintendo 64DD","Nintendo DS", "Nintendo DSi","Super NES","Wii","Wii U")

sony <- c("PlayStation","PlayStation 2","PlayStation 3","PlayStation 4" ,"PlayStation Portable","PlayStation Vita")

microsoft <- c("Xbox","Xbox 360","Xbox One")

# Function to create new console platform
consolePlatform <- function(x) {
  if (x %in% nintendo == TRUE){
    return("Nintendo")
  }
  
  else if(x %in% sony == TRUE){
    return("Sony")  
  }
  
  else if(x %in% microsoft == TRUE){
    return("Microsoft")
  }
  
  else{
    return("Others")
  }
  
}

ign20$consolePlatform <- sapply(ign20$platform, consolePlatform)
table(ign20$consolePlatform)
```
Sony won total released game on their platform with total 5059
Nintendo in 2nd position with 4907 total games
and Microsoft in 3rd position with 2659 total games

## 6. Data Visualization
### 6.1 Nintendo 1996-2016
```{r}
ggplot(ign20[ign20$consolePlatform=="Nintendo",],aes(x=factor(release_year))) +
  geom_bar(aes(fill=platform)) +
  theme(axis.text.x = element_text(angle=90, hjust=1)) +
  labs(title = "Total game were released every year in Nintendo console", x = "Release Year", y = "Total Game")

summary(ign20[ign20$consolePlatform=="Nintendo",])
```
Nintendo was the most senior game company compare to Sony and Microsoft.
Nintendo has the worst year during 1996-1998, because Sony almost take down all the console game marketshare with their premiere platform.
With the release of Wii, Nintendo is rising again with their most innovative game console
Unfortunately, the Wii successor, Wii U, didn't work well as its predecessor, and make Nintendo has the worst year again during 2014-2016.


### 6.2 Sony 1996-2016
```{r}
ggplot(ign20[ign20$consolePlatform=="Sony",],aes(x=factor(release_year))) +
  geom_bar(aes(fill=platform)) +
  theme(axis.text.x = element_text(angle=90, hjust=1)) +
  labs(title = "Total game were released every year in Sony console", x = "Release Year", y = "Total Game")

summary(ign20[ign20$consolePlatform=="Sony",])
```
Sony has been active in console game industry since 1996 with the release of phenomenal PlayStation 1.
Until 2016, PlaySation 2 still the most popular console game for game release.


## 6.3 Microsoft 1996-2016
```{r}
ggplot(ign20[ign20$consolePlatform=="Microsoft",],aes(x=factor(release_year))) +
  geom_bar(aes(fill=platform)) +
  theme(axis.text.x = element_text(angle=90, hjust=1)) +
  labs(title = "Total game were released every year in Microsoft console", x = "Release Year", y = "Total Game")

summary(ign20[ign20$consolePlatform=="Microsoft",])
```
We can see that Microsoft doesn't have any game released before 2001, that was because the first Microsoft game console, Xbox, was release on 2001. That means Bill Gates own-company doesn't compete with Nintendo and Sony before that time.

Xbox 360 was the most popular Microsoft game platform

## 6.4 Console Wars
What is the most popular console during 1996-2016
```{r}
a <- ggplot(ign20[ign20$consolePlatform != "Others",],aes(x=factor(release_year))) +
  geom_bar(aes(fill=consolePlatform)) +
  theme(axis.text.x = element_text(angle=90, hjust=1)) +
  labs(title = "Total game were released every year", x = "Release Year", y = "Total Game")

library(plotly)
ggplotly(a)

```


## 7. Final Thoughts
Sony wins the 20-years **Console Wars** based on their consistency in game release every year, with more than 5000 game title released on their platform from 1996-2016.


## 8. What do I learn?
* Pre-processing data by subsetting date based on **release_year** column from ign dataframe
* Data preparation by creating new variable **consolePlatform** using 3 big console game vendor: Nintendo, Sony, and Microsoft
* Data visualization with ggplot2 using `geom_bar()` + `theme()` + `labs()`
* Analyze 20 years **Console Wars** with timeframe 1996-2016

## References
[kaggle.com](https://www.kaggle.com/jonathanbouchet/ign-data-20-years-of-vg)
