---
title: "R Notebook"
output: html_notebook
---


insample accuracy
accuracy model


QUIZ
The sales director at your bank in charge of telemarketing and direct marketing approach you for help:

Instead of ringing up every customer in the database only to be rejected more than 88% of the time, she would like you to use the age, education, job profile, existing loan records (housing / personal) and outcome of the previous campaign to predict the customer's likelihood to subscribe to the new financial product (term deposit).

The model output a probability (tip: bankmar_nn$net.result) so you'll need to convert that to a 'yes' or 'no'. The ground truth label is stored in your original data bank$y.

```{r}
prop.table(table(bank$y)) 
```

Question 1
What is the accuracy of your model (in percentage to 2 decimal point, ie: 75.13, 80.21, etc.)? 
`89.34`


Question 2
The accuracy isn't very meaningful in this case. Even if your model were to predict using a One-Rule approach and classified all the cases as 'no', you'd still obtain an 88.3% accuracy. Instead, the primary objective based on your colleague's requirement is to output a list from the initial 45,211 customers that have a high likelihood to purchase the term deposit through her telemarketing team. How long is the list of high-potential leads you've identified? 
`1461`


Question 3
Out of the customers you've identified as 'high target', how many of them did actually sign up for the term deposit (in percentage to 2 decimal point, ie: 75.13, 80.21, etc.)? `66.05`

Question 4
What you just performed above in question (3) is also known as: 
A.  Precision 
B.  Recall 
C.  Specificity 
`Precision`


```{r}
bankmar_nn <- readRDS(file = "bank_marketing_nn.rds")
```

```{r}
library(neuralnet)
length(bankmar_nn$net.result)
pred <- bankmar_nn$net.result[[1]]
# summary(pred)
pred <- factor(round(pred), levels=c(0,1), labels = c("no","yes"))
```

```{r}
library(caret)
confusionMatrix(pred, bank$y, positive = "yes")
```

